Opening up the court (surface) in tennis grand slams

### 2019-09-05 paper_results/Data/data-description.Rmd
Updated data description
re-organize and clean files

---
title: "Data description"
author: "Gallagher, Frisoli, and Luby"
date: "September 5, 2019"
output:
  html_document
---

```{r setup, eval = FALSE}
library(courtsports)
library(knitr)
library(kableExtra)
library(broom)
library(tidyverse)
```

```{r invis-setup, echo = FALSE, warning = FALSE, message = FALSE}
devtools::load_all("~/courtsports")
library(knitr)
library(kableExtra)
library(broom)
library(tidyverse)
```

>>>>

---
title: "Data description"
author: ""
date: "September 5, 2019"
output:
  html_document
---

```{r invis-setup, echo = FALSE, warning = FALSE, message = FALSE}
# devtools::install_github("shannong19/courtsports")
library(courtsports)
library(knitr)
library(kableExtra)
library(broom)
library(tidyverse)
```

### 2019-09-05 paper_results/EDA/eda.Rmd
small edits
re-organize and clean files

---
title: "EDA"
author: ""
date: "Sept. 5, 2019"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding: hide
---

@@

```{r libraries}
library(devtools)
library(deuce)
library(knitr)
#library(courtsports) ## UNCOMMENT if running
#devtools::load_all("..") # Loads courtsports package

>>>>

---
title: "EDA"
author: ""
date: " "
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding: hide
---

@@

```{r libraries}
library(devtools)
# devtools::install_github("skoval/deuce")
library(deuce)
library(knitr)
# devtools::install_github("shannong19/courtsports")
library(courtsports)
```

### 2019-09-05 paper_results/EDA/graphics_eda.Rmd
re-organize and clean files

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message = FALSE)
library(deuce)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(grid)
library(MASS)
library(ggrepel)
library(broom)
library(lme4)
library(cowplot)
library(ggpubr)
library(reshape2)

## To change into library(courtsports)
devtools::load_all("../..")
```

@@

tournament_colors <- c("#0297DB", "#b06835", "#0C2340", "#54008b")
league_colors <- c("#0C2340", "#902BA3")
win_colors <- c("#336699", "#339966")

# with yellow for us
tournament_colors <- c("#0297DB", "#b06835", "#ffe500", "#54008b")

@@

if(save_graph){
  grid.arrange(p2, p1, ncol=2) %>%
  ggsave("plots/points_per_match.jpg", ., height=8, width=graphic_width_long)
}
```

@@

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) 
if(save_graph){
ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court.jpg", ., height=5.5, width=graphic_width_long)
}
```

@@

if(save_graph){
ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court_all4.jpg", ., height=5, width=graphic_width_long)
}
```

@@

if(save_graph){
ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
  ggsave("plots/spainards.jpg", ., height=6, width=graphic_width_long)
}

@@

if(save_graph){
ggsave("plots/errors_aces2.jpg", height = 6, width = graphic_width_short)
}
```

@@

if(save_graph){
ggsave("plots/aces2.jpg", height=7, width=graphic_width_long)
}
```

@@

if(save_graph){
ggsave("plots/aces2_percent.jpg", height=8, width=graphic_width_long)
}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message = FALSE)

# devtools::install_github("skoval/deuce")
library(deuce)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(grid)
library(MASS)
library(ggrepel)
library(broom)
library(lme4)
library(cowplot)
library(ggpubr)
library(reshape2)
# devtools::install_github("shannong19/courtsports")
library(courtsports)
```

@@

league_colors <- c("#0C2340", "#902BA3")
win_colors <- c("#336699", "#339966")
# with yellow for USA
tournament_colors <- c("#0297DB", "#b06835", "#ffe500", "#54008b")

@@

if(save_graph){
  grid.arrange(p2, p1, ncol=2) %>%
  ggsave("../plots/points_per_match.jpg", ., height=8, width=graphic_width_long)
}
```

@@

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) 
if(save_graph){
ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("../plots/home_court.jpg", ., height=5.5, width=graphic_width_long)
}
```

@@

if(save_graph){
ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("../plots/home_court_all4.jpg", ., height=5, width=graphic_width_long)
}
```

@@

if(save_graph){
ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
  ggsave("../plots/spainards.jpg", ., height=6, width=graphic_width_long)
}

@@

if(save_graph){
ggsave("../plots/errors_aces2.jpg", height = 6, width = graphic_width_short)
}
```

@@

if(save_graph){
ggsave("../plots/aces2.jpg", height=7, width=graphic_width_long)
}
```

@@

if(save_graph){
ggsave("../plots/aces2_percent.jpg", height=8, width=graphic_width_long)
}
```

### 2019-09-05 paper_results/Hierarchical/graphics_model_complete_data.Rmd
re-organize and clean files

tournament_colors <- c("#0297DB", "#b06835", "#0C2340", "#54008b")
league_colors <- c("#0C2340", "#902BA3")
win_colors <- c("#336699", "#339966")

# with yellow for us
tournament_colors <- c("#0297DB", "#b06835", "#ffe500", "#54008b")

@@

if(save_graph){
ggarrange(aces_ranef, nets_ranef, ue_ranef, nrow=1, ncol=3,
          common.legend = TRUE, legend="bottom", widths = c(.42, .29, .29)) %>%
ggsave("plots/player_effects.jpg", ., height=6, width = graphic_width_long)
}
```

@@

if(save_graph){
ggsave("plots/ref-courts-eff.jpg", height=6, width = graphic_width_short)
}
```

@@

```{r, fig.height=7, fig.width=16, fig.align='center'}
win_colors <- c("#336699", "#339966")

model_names = c("nocourt_logistic", "base_logistic", "country_logistic", "ind_logistic", 
                "ind_logistic_noioc", "ind_int_logistic", "ind_year_logistic")

model.sums = data.frame(model = model_names, 
                        aic = rep(NA, length(model_names)), 
                        edf = rep(NA,length(model_names)))
model.ests = data.frame()
ran.effects = data.frame()
for(mod in 1:nrow(model.sums)){
  name = model_names[mod]
  model.sums$aic[mod] = extractAIC(get(name))[2]
  model.sums$edf[mod] = extractAIC(get(name))[1]
  betas = tidy(get(name))
  betas$model = name
  if("group" %in% colnames(betas)){
    model.ests = rbind(model.ests, filter(betas, group == "fixed") %>%
                         dplyr::select(., -c("group")))
  }
  else model.ests = rbind(model.ests, betas)
}

model.sums$name = c("no_court", "no_random_ef", "country_ef", "ind_ef",
                    "ind_no_ioc", "ind_intercept", "ind_year_ef")
model.sums$fixed = c("No country, no tournament",
                     "all",
                     "no country, no tournament",
                     "no tournament",
                     "no country, no tournament",
                     "no country",
                     "no country, no year")
model.sums$random = c("none",
                      "none",
                      "tournament (by country)",
                      "tournament (by individual)",
                      "tournament (by individual)",
                      "intercept (by individual)",
                      "year (by individual)")

fixed.effects.plot = model.ests %>%
  filter(., !grepl("ioc", term)) %>% 
  filter(., model == 'ind_logistic_noioc') %>%
  ggplot(., aes(x = term,
                y = estimate,
                col = model,
                ymin = estimate - 2*std.error,
                ymax = estimate + 2*std.error)) +
  geom_errorbar(width = .7, position = position_dodge(), size = 1.5) +
  my_theme +
  theme(axis.text.x=element_text(angle = 35, hjust = 1)) +
  ggtitle("Fixed effects") +
  scale_color_brewer("",palette = "Dark2") +
  theme(legend.position = "none") + 
  xlab("") +
  ylab("Coefficient") +
  scale_x_discrete("", breaks = waiver(), 
                   labels = c("Intercept", "ATP", "Late round", "Op Rank (log)",
                              "Rank (log)", "2014", "2015", "2016", "2017")) +
  guides(col = guide_legend(nrow=2)) +
  geom_hline(yintercept = 0, color = "grey") + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

ran_tourn = rbind(ranef(ind_logistic)$name_int, ranef(ind_logistic_noioc)$name_int)
ran_tourn$model = c(rep("ind_logistic", 
                        nrow(ranef(ind_logistic)$name_int)), 
                    rep("ind_logistic_noioc",
                        nrow(ranef(ind_logistic)$name_int)))

gs_players$name_int = as.integer(as.factor(gs_players$name))
name_lookup = unique(cbind(gs_players$name, gs_players$name_int))
name_lookup = name_lookup[order(as.integer(name_lookup[,2])), ]

top_three_ranef = as.data.frame(ranef(ind_logistic_noioc, condVar = TRUE)) %>%
  mutate(tournament = gsub("tournament", "", term)) %>%
  mutate(name = name_lookup[,1][as.integer(grp)]) %>% 
  filter(., name == "Serena Williams" | name == "Rafael Nadal" | name == "Roger Federer") %>%
  ggplot(., aes(x = name, 
                y = condval, 
                ymin = condval- 2*condsd, 
                ymax = condval + 2*condsd, 
                color = tournament)) +
  geom_errorbar(width = .5, position = "dodge",size = 1.5) + 
  xlab("") + 
  ylab("Coefficient") + 
  scale_color_manual("", values=tournament_colors) +
  my_theme +
  coord_flip() +
  ggtitle("Random Effects") +
  theme(legend.position = "bottom") + 
  guides(col = guide_legend(nrow=2,
                            override.aes = list(size = 7))) +
  geom_hline(yintercept = 0, color = "grey") + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

grid.arrange(fixed.effects.plot, top_three_ranef, nrow = 1, widths = c(.45, .55))

if(save_graph){
grid.arrange(fixed.effects.plot, top_three_ranef, nrow = 1, widths = c(.45, .55)) %>%
ggsave("plots/logistic-parameter-effects.jpg",., height=7, width=graphic_width_long)
}
```

@@

if(save_graph){
ggsave("plots/linear-fixed-effects.jpg", fixed.effects.plot.linear,
       height=6, width = graphic_width_short)
}
```

@@

if(save_graph){
ggsave("plots/correlation.jpg", 
       height=6, width = graphic_width_short)
}
```

>>>>

league_colors <- c("#0C2340", "#902BA3")
win_colors <- c("#336699", "#339966")
# with yellow for us
tournament_colors <- c("#0297DB", "#b06835", "#ffe500", "#54008b")

@@

if(save_graph){
ggarrange(aces_ranef, nets_ranef, ue_ranef, nrow=1, ncol=3,
          common.legend = TRUE, legend="bottom", widths = c(.42, .29, .29)) %>%
ggsave("../plots/player_effects.jpg", ., height=6, width = graphic_width_long)
}
```

@@

if(save_graph){
ggsave("../plots/ref-courts-eff.jpg", height=6, width = graphic_width_short)
}
```

@@

```{r, fig.height=7, fig.width=16, fig.align='center'}
# win_colors <- c("#336699", "#339966")
# 
# model_names = c("nocourt_logistic", "base_logistic", "country_logistic", "ind_logistic", 
#                 "ind_logistic_noioc", "ind_int_logistic", "ind_year_logistic")
# 
# model.sums = data.frame(model = model_names, 
#                         aic = rep(NA, length(model_names)), 
#                         edf = rep(NA,length(model_names)))
# model.ests = data.frame()
# ran.effects = data.frame()
# for(mod in 1:nrow(model.sums)){
#   our_name = model_names[mod]
#   model.sums$aic[mod] = extractAIC(get(our_name))[2]
#   model.sums$edf[mod] = extractAIC(get(our_name))[1]
#   betas = tidy(get(our_name))
#   betas$model = our_name
#   if("group" %in% colnames(betas)){
#     model.ests = rbind(model.ests, filter(betas, group == "fixed") %>%
#                          dplyr::select(., -c("group")))
#   } else model.ests = rbind(model.ests, betas)
# }
# 
# model.sums$our_name = c("no_court", "no_random_ef", "country_ef", "ind_ef",
#                     "ind_no_ioc", "ind_intercept", "ind_year_ef")
# model.sums$fixed = c("No country, no tournament",
#                      "all",
#                      "no country, no tournament",
#                      "no tournament",
#                      "no country, no tournament",
#                      "no country",
#                      "no country, no year")
# model.sums$random = c("none",
#                       "none",
#                       "tournament (by country)",
#                       "tournament (by individual)",
#                       "tournament (by individual)",
#                       "intercept (by individual)",
#                       "year (by individual)")
# 
# fixed.effects.plot = model.ests %>%
#   filter(., !grepl("ioc", term)) %>% 
#   filter(., model == 'ind_logistic_noioc') %>%
#   ggplot(., aes(x = term,
#                 y = estimate,
#                 col = model,
#                 ymin = estimate - 2*std.error,
#                 ymax = estimate + 2*std.error)) +
#   geom_errorbar(width = .7, position = position_dodge(), size = 1.5) +
#   my_theme +
#   theme(axis.text.x=element_text(angle = 35, hjust = 1)) +
#   ggtitle("Fixed effects") +
#   scale_color_brewer("",palette = "Dark2") +
#   theme(legend.position = "none") + 
#   xlab("") +
#   ylab("Coefficient") +
#   scale_x_discrete("", breaks = waiver(), 
#                    labels = c("Intercept", "ATP", "Late round", "Op Rank (log)",
#                               "Rank (log)", "2014", "2015", "2016", "2017")) +
#   guides(col = guide_legend(nrow=2)) +
#   geom_hline(yintercept = 0, color = "grey") + 
#   theme(panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank())
# 
# ran_tourn = rbind(ranef(ind_logistic)$name_int, ranef(ind_logistic_noioc)$name_int)
# ran_tourn$model = c(rep("ind_logistic", 
#                         nrow(ranef(ind_logistic)$name_int)), 
#                     rep("ind_logistic_noioc",
#                         nrow(ranef(ind_logistic)$name_int)))
# 
# gs_players$name_int = as.integer(as.factor(gs_players$our_name))
# name_lookup = unique(cbind(gs_players$name, gs_players$name_int))
# name_lookup = name_lookup[order(as.integer(name_lookup[,2])), ]
# 
# top_three_ranef = as.data.frame(ranef(ind_logistic_noioc, condVar = TRUE)) %>%
#   mutate(tournament = gsub("tournament", "", term)) %>%
#   mutate(our_name = name_lookup[,1][as.integer(grp)]) %>% 
#   filter(., our_name == "Serena Williams" | our_name == "Rafael Nadal" | our_name == "Roger Federer") %>%
#   ggplot(., aes(x = our_name, 
#                 y = condval, 
#                 ymin = condval- 2*condsd, 
#                 ymax = condval + 2*condsd, 
#                 color = tournament)) +
#   geom_errorbar(width = .5, position = "dodge",size = 1.5) + 
#   xlab("") + 
#   ylab("Coefficient") + 
#   scale_color_manual("", values=tournament_colors) +
#   my_theme +
#   coord_flip() +
#   ggtitle("Random Effects") +
#   theme(legend.position = "bottom") + 
#   guides(col = guide_legend(nrow=2,
#                             override.aes = list(size = 7))) +
#   geom_hline(yintercept = 0, color = "grey") + 
#   theme(panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank())
# 
# grid.arrange(fixed.effects.plot, top_three_ranef, nrow = 1, widths = c(.45, .55))
# 
# if(save_graph){
# grid.arrange(fixed.effects.plot, top_three_ranef, nrow = 1, widths = c(.45, .55)) %>%
# ggsave("../plots/logistic-parameter-effects.jpg",., height=7, width=graphic_width_long)
# }
```

@@

if(save_graph){
ggsave("../plots/linear-fixed-effects.jpg", fixed.effects.plot.linear,
       height=6, width = graphic_width_short)
}
```

@@

if(save_graph){
ggsave("../plots/correlation.jpg", 
       height=6, width = graphic_width_short)
}
```
